Cradles of the first stars: self-shielding, halo masses, and multiplicity

### 2019-12-05 paper.tex
Changes after my 1st proofread

If within \textbf{10} pc, multiple cells meet this criteria, then a single Pop III star forms at the center of mass of these cells. In a parameter study, we have found that the number of Pop III star particles is insensitive to this algorithmic merging distance. \textbf{We found that even when this number was decreased to 0.065 pc, the number of star particles that formed was unchanged.  Our density threshold for star particle formation is equivalent to a molecular core.  Fragmentation of metal-free gas into massive stellar precursors can occur at scales down to 1000~AU \citep[e.g.][]{Turk09, Stacy11} with molecular cores, therefore our results on multiplicity should be considered as lower limits.}

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/jlw_xe_bg_mass_weight.pdf}
    \caption{For the entire box, the average mass weighted LW background from sources shown in solid red and the fraction of the volume ionized above 10\% versus time shown in solid black. The red dashed line represents the LW background implemented in the simulation (see Equation \ref{LWbg}). \textbf{The solid red line in the legend refers to both the solid and dashed lines in the figure.}Temperature projections are shown at the points indicated by the arrows.}
    \label{fig:JLW_xe_mass}
\end{figure}

In our simulation, 688 Pop III stars form between 27.23 $\geq z \geq$ 9.39. The corresponding \textbf{comoving} star formation rate density (SFRD) can be seen in Figure \ref{fig:pop3_SFR_bar}, peaking at $1.9 \times 10^{-4} \ \rm M_{\odot} \ yr^{-1} \ Mpc^{-3}$ at $z = 21.3$ and slowly decreases. The SFRD sharply declines near the end of the simulation but does not drop to zero. Small volumes are subject to cosmic variance, in particular the formation time of the first galaxies, which generates the bulk of the LW radiation background.  This SFRD is therefore not representative of a cosmological average.  In particular, it decreases at $z \le 11$ when the most massive galaxy undergoes an extended period of star formation, boosting the local LW radiation field as shown in Figure \ref{fig:JLW_xe_mass}. \textbf{\citet{Xu13} and \citet{Magg16} see a SFRD of $~10^{-4} \rm M_{\odot} \ yr^{-1} \ Mpc^{-3}$ at $z = 15$ and $z = 20$ respectively. \citet{Jaacks19} find a SFRD of $~10^{4.5} \rm M_{\odot} \ yr^{-1} \ Mpc^{-3}$ at $z = 20$, which flattens out to $10^{-3} \rm M_{\odot} \ yr^{-1} \ Mpc^{-3}$ by $z =7$.} \textbf{Our} SFRD rates are consistent with \textbf{these} other studies, but it is likely to change depending on the particular modes captured by the initial conditions.

@@

We now inspect the number of Pop III stars and the total mass of Pop III stars per halo. Again, we are only tracking \textbf{star-forming regions where the star particles form at densities equivalent to molecular cores that may form multiple massive stars.  Furthermore, we are not modeling the formation of low-mass stars ($<1 M_{\odot}$) that may form out of fragmented gas at higher densities and smaller scales.  Thus our results in this section should be considered as lower limits to the number of Pop III stars.}  The left panel of Figure \ref{fig:num_p3} shows the number of Pop III stars per halo for a given halo mass, just after star formation. The histogram of the number of Pop III stars for all masses is projected on the right hand side. We find that a median number of four Pop III stars form per halo, with a maximum of 16 Pop III stars forming per halo. Since we did not restrict the number of Pop III stars that can form in a halo, we find that the conditions are often sufficient for multiple Pop III stars to form in a single halo. Out of the halos forming Pop III stars, only 16\% of them form single Pop III stars, whereas 54\% form between two and five Pop III stars, similar to the radiation hydrodynamic simulations of \citet{Susa13, Susa14}. It should be noted that these halos are not necessarily forming all of the Pop III stars at once. For example, a halo can form stars again if it did not host a supernova. The number of Pop III stars that form within a halo is indicative of the star formation history of that halo. Figure \ref{fig:totp3mass_halomass_sidehist} shows the total mass of Pop III stars per halo for a given halo mass, where the histogram of the total Pop III mass for all halos is projected on the right hand side. Lines of constant star formation efficiency are overplotted. We find that most of our Pop III stars are forming between 10$^{-4}$ < f$_\star$ < 10$^{-3}$. The mean total mass of Pop III stars for all halos is 195 M$_{\odot}$. Note that the mass of the Pop III stars depends on the chosen IMF.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/compare_JLW_mass.pdf}
    \caption{The halo mass versus the average LW intensity normalized by J21 versus for this work and a variety of other works. The black solid line is from M01 given our LW background, blue solid line from \citet{Trenti09_SFR} at $z=20$, the purple solid line from \citet{Visbal14} at $z=20$, blue open circle from \citet{Yoshida03}, green open squares from \citet{OShea08}, and red open triangles from \citet{Wise07_UVB}. J$_{LW}$ = 0 for the various sources is plotted here at J$_{\rm LW}$/J$_{21}$ = 10$^{-4}$. \textbf{The left panel represents halo masses corresponding to different streaming velocities from \citet{Naoz13} and \citet{Schauer19}.}The solid bands of halo masses correspond to the characteristic masses of halos to contain a baryon fraction that is half of the cosmic mean for various redshifts and for the shown streaming velocities from \citet{Naoz13}. For each band, the maximum halo mass occurs at $z = 15$. For streaming velocities of $3.4 \sigma_{\rm vbc}$, $1.7 \sigma_{\rm vbc}$, and $1 \sigma_{\rm vbc}$, the minimum halo mass occurs at $z = 17, 22, \rm{and} \ 27$, respectively. \textbf{The minimum halo mass hosting cold dense gas for different streaming velocities from \citet{Schauer19} are shown as horizontal black lines.}}
    \label{fig:compare_JLW_mass}
\end{figure}

@@

There are a few minor shortcomings to this work that should be noted. \textbf{As discussed in section 2.2.1, the number of Pop III stars that form is insensitive down to distances of 0.625 pc. Since this value is larger than fragmentation scales, it's possible that the location of the Pop III stars formed in our simulation are sites for multiple Pop III star formation. With this in mind, our results are only strengthened by the possibility of having more Pop III stars forming at these locations.}

@@

We also do not include streaming velocities between baryons and dark matter within our simulation \citep{Tselia11, Greif11_Delay, Naoz12, OLeary12, Schauer19, Hirano17_Science}. Streaming velocities can suppress star formation at very high ($z \ga 20$) redshifts only, because the streaming velocities decrease with time. Pop III star formation is delayed by an average of $\delta z = 4$ by increasing the halo mass needed to overcome the bulk velocity by about a factor of three \citep{Greif11_Delay}. The streaming velocities will generally result in lower gas densities within halos as well as an offset of the peak density from the center of the halo \citep{OLeary12}. \citet{Naoz13} estimated the minimum mass needed to retain the bulk of the baryons within the dark matter halo using a series of cosmological simulations, the results of which can be seen in Figure \ref{fig:compare_JLW_mass} (see shaded bands). These halos only reach up to about $2 - 3 \times 10^5 M_\odot$ when the streaming velocity is about $1\sigma_{\rm vbc}$, which is smaller than the star forming halos in our simulation. \citet{Naoz13} mention that at the largest streaming velocity, their simulations do not match expected values due to a small sample size of high mass halos.

\textbf{A more recent study by \citet{Schauer19}, who used cosmological simulations to study cold gas content within halos for varying streaming velocities, showed that high streaming velocities restrict first star formation to atomic cooling halos. Their minimum halo masses which host cold dense gas for various streaming velocities are shown as black horizontal lines with different linestyles in Figure \ref{fig:compare_JLW_mass}. \citet{Hirano17_Science} investigated the effect that streaming velocities have on black hole formation in the early universe and find that in regions with high streaming velocities and for z $> 30$, first generation star formation can be suppressed in minihalos up to the atomic cooling limit, and may produce massive black hole seeds.}

Our results indicate that halos collapse at larger masses that are not as affected by typical streaming velocities ($<1\sigma_{\rm vbc}$). \textbf{We also find that Pop III stars begin to form at much later times than the halos found in \citet{Hirano17_Science}.} In the evolution of such a halo, it would have started to accumulate gas above these critical masses while also being irradiated by a LW background.  Afterwards, the collapse will be controlled by the processes discussed in Section \ref{sec:variations}, and thus streaming velocities would likely not affect our results.

>>>>

If within \textbf{10} pc, multiple cells meet this criteria, then a single Pop III star forms at the center of mass of these cells. In a parameter study, we have found that the number of Pop III star particles is insensitive to this algorithmic merging distance. \textbf{We ran another 1~Mpc$^3$ simulation with the same setup as the main simulation with the exception that it was a zoom-in simulation focusing on a single minihalo of mass $5 \times 10^5~\Ms$ collapsing at $z \simeq 23$.  Starting from 1~Myr before the collapse, we followed Pop III star formation in this one halo with a merging distance of 10, 5, 2.5, 1.25, and 0.625~pc.  We found that even when this number was decreased to 0.065 pc, the number of star particles that formed was unchanged.  We note that our density threshold for star particle formation is equivalent to a molecular core.  Fragmentation of metal-free gas into massive stellar precursors can occur at scales down to 1000~AU \citep[e.g.][]{Turk09, Stacy11} with molecular cores, therefore our results on multiplicity should be considered as lower limits.}

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/jlw_xe_bg_mass_weight.pdf}
    \caption{For the entire box, the average mass weighted LW background from sources shown in solid red and the fraction of the volume ionized above 10\% versus time shown in solid black. The red dashed line represents the LW background implemented in the simulation (see Equation \ref{LWbg}). \textbf{The solid red line in the legend refers to both the solid and dashed lines in the figure.} Temperature projections are shown at the points indicated by the arrows.}
    \label{fig:JLW_xe_mass}
\end{figure}

In our simulation, 688 Pop III stars form between 27.23 $\geq z \geq$ 9.39. The corresponding \textbf{comoving} star formation rate density (SFRD) can be seen in Figure \ref{fig:pop3_SFR_bar}, peaking at $1.9 \times 10^{-4} \ \rm M_{\odot} \ yr^{-1} \ Mpc^{-3}$ at $z = 21.3$ and slowly decreases. The SFRD sharply declines near the end of the simulation but does not drop to zero. Small volumes are subject to cosmic variance, in particular the formation time of the first galaxies, which generates the bulk of the LW radiation background.  This SFRD is therefore not representative of a cosmological average.  In particular, it decreases at $z \le 11$ when the most massive galaxy undergoes an extended period of star formation, boosting the local LW radiation field as shown in Figure \ref{fig:JLW_xe_mass}. \textbf{\citet{Xu13} and \citet{Magg16} see a SFRD of $~10^{-4}~\rm M_{\odot} \ yr^{-1} \ Mpc^{-3}$ at $z = 15$ and $z = 20$ respectively. \citet{Jaacks19} find a SFRD of $~10^{4.5}~\rm M_{\odot} \ yr^{-1} \ Mpc^{-3}$ at $z = 20$, which flattens out to $10^{-3}~\rm M_{\odot} \ yr^{-1} \ Mpc^{-3}$ by $z =7$.} \textbf{Our} SFRD rates are consistent with \textbf{these} other studies, but it is likely to change depending on the particular modes captured by the initial conditions.

@@

We now inspect the number of Pop III stars and the total mass of Pop III stars per halo. We are only tracking \textbf{star-forming regions where the star particles form at densities equivalent to molecular cores that may form multiple massive stars.  Furthermore, we are not modeling the formation of low-mass stars ($<1 M_{\odot}$) that may form out of fragmented gas at higher densities and smaller scales.  Thus our results in this section should be considered as lower limits to the number of Pop III stars.}  The left panel of Figure \ref{fig:num_p3} shows the number of Pop III stars per halo for a given halo mass, just after star formation. The histogram of the number of Pop III stars for all masses is projected on the right hand side. We find that a median number of four Pop III stars form per halo, with a maximum of 16 Pop III stars forming per halo. Since we did not restrict the number of Pop III stars that can form in a halo, we find that the conditions are often sufficient for multiple Pop III stars to form in a single halo. Out of the halos forming Pop III stars, only 16\% of them form single Pop III stars, whereas 54\% form between two and five Pop III stars, similar to the radiation hydrodynamic simulations of \citet{Susa13, Susa14}. It should be noted that these halos are not necessarily forming all of the Pop III stars at once. For example, a halo can form stars again if it did not host a supernova. The number of Pop III stars that form within a halo is indicative of the star formation history of that halo. Figure \ref{fig:totp3mass_halomass_sidehist} shows the total mass of Pop III stars per halo for a given halo mass, where the histogram of the total Pop III mass for all halos is projected on the right hand side. Lines of constant star formation efficiency are overplotted. We find that most of our Pop III stars are forming between 10$^{-4}$ < f$_\star$ < 10$^{-3}$. The mean total mass of Pop III stars for all halos is 195 M$_{\odot}$. Note that the mass of the Pop III stars depends on the chosen IMF.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{images/compare_JLW_mass.pdf}
    \caption{The halo mass versus the average LW intensity normalized by J21 versus for this work and a variety of other works. The black solid line is from M01 given our LW background, blue solid line from \citet{Trenti09_SFR} at $z=20$, the purple solid line from \citet{Visbal14} at $z=20$, blue open circle from \citet{Yoshida03}, green open squares from \citet{OShea08}, and red open triangles from \citet{Wise07_UVB}. J$_{LW}$ = 0 for the various sources is plotted here at J$_{\rm LW}$/J$_{21}$ = 10$^{-4}$. \textbf{The right panel represents halo masses corresponding to different streaming velocities from \citet{Naoz13} and \citet{Schauer19}} The solid bands of halo masses correspond to the characteristic masses of halos to contain a baryon fraction that is half of the cosmic mean for various redshifts and for the shown streaming velocities from \citet{Naoz13}. For each band, the maximum halo mass occurs at $z = 15$. For streaming velocities of $3.4 \sigma_{\rm vbc}$, $1.7 \sigma_{\rm vbc}$, and $1 \sigma_{\rm vbc}$ (5.8 km/s at $z=199$), the minimum halo mass occurs at $z = 17, 22, \rm{and} \ 27$, respectively. \textbf{The minimum halo mass hosting cold dense gas for different streaming velocities, given at their starting redshift $z=199$, from \citet{Schauer19} are shown as horizontal black lines.}}
    \label{fig:compare_JLW_mass}
\end{figure}

@@

There are a few minor shortcomings to this work that should be noted. \textbf{As discussed in \S2.2.1, the number of Pop III stars that form is insensitive down to distances of 0.625 pc. Since this value is larger than fragmentation scales, it is possible that the location of the Pop III star particles formed in our simulation are sites for multiple Pop III star formation. With this in mind, our multiplicity results are only strengthened by the possibility of having more Pop III stars forming at these locations.}

@@

We also do not include streaming velocities between baryons and dark matter within our simulation \citep{Tselia11, Greif11_Delay, Naoz12, OLeary12, Schauer19, Hirano17_Science}. Streaming velocities can suppress star formation at very high ($z \ga 20$) redshifts only, because the streaming velocities decrease inversely with the scale factor. Pop III star formation is delayed by an average of $\delta z = 4$ by increasing the halo mass needed to overcome the bulk velocity by about a factor of three \citep{Greif11_Delay}. The streaming velocities will generally result in lower gas densities within halos as well as an offset of the peak density from the center of the halo \citep{OLeary12}. \citet{Naoz13} estimated the minimum mass needed to retain the bulk of the baryons within the dark matter halo using a series of cosmological simulations, the results of which can be seen in Figure \ref{fig:compare_JLW_mass} (see shaded bands). These halos only reach up to about $2 - 3 \times 10^5 M_\odot$ when the streaming velocity is about $1\sigma_{\rm vbc} = 3$ km/s at $z=99$, which is smaller than the star forming halos in our simulation. \citet{Naoz13} mention that at the largest streaming velocity, their simulations do not match expected values due to a small sample size of high mass halos.

\textbf{A more recent study by \citet{Schauer19}, who used cosmological simulations to study cold gas content within halos for varying streaming velocities, showed that high streaming velocities could restrict first star formation to atomic cooling halos in the most extreme cases. Their minimum halo masses which host cold dense gas for various streaming velocities are shown as black horizontal lines in Figure \ref{fig:compare_JLW_mass}. \citet{Hirano17_Science} investigated the effect that streaming velocities have on massive black hole seed formation and find that in regions with high streaming velocities and for z $> 30$, first generation star formation can be suppressed in minihalos up to the atomic cooling limit.}

\textbf{The results of \citet{Naoz13} and \citep{Schauer19} in Figure \ref{fig:compare_JLW_mass} show that typical streaming velocities ($\le 1\sigma_{\rm vbc} = 5.8$~km/s at $z=199$) would affect halos below $10^6~\Ms$, especially at redshifts $z>20$ when streaming velocities are the strongest.  Our results sample minihalos that experience such streaming velocities primarily at $z < 20$, whereas \citet{Hirano17_Science} studied much earlier structure formation. In the evolution of a halo affected by streaming velocities, it would have started to accumulate gas once above these critical masses while also being irradiated by a LW background.  Afterwards, the collapse will be controlled by the processes discussed in Section \ref{sec:variations}, and thus streaming velocities would likely not affect our main conclusion that \hh{} plays an important role in Pop III star formation.}
